# Amplification Circuit

[Instructions](https://adventofcode.com/2019/day/7)

```{r}
library(gtools)

read_day7 <- function() {
  raw <- readLines("data-raw/day07.txt")
  vec <- strsplit(raw, ",")[[1]]
  as.numeric(vec)
}
```

## Part 1

```{r}
posib <- gtools::permutations(
  n = 5, r = 5, v = 0:4, 
  repeats.allowed=FALSE
  )
  
run_intcode <- function(phase_setting, input_signal = 0) {
  
  running_state <- read_day7()
  
  input <- c(phase_setting, input_signal)
  instr_index <- 1
  
  while (instr_index < length(running_state)) {
    
    # Instruction & parameters
    
    instr_orig <- running_state[instr_index]
    instr_full <- stringr::str_pad(instr_orig, 5, "left", "0")
    
    instr_params <- stringr::str_split(instr_full, "")[[1]]
    
    optcode <- instr_params[5]
    param1  <- instr_params[3]
    param2  <- instr_params[2]
    
    
    # Execute the program
    
    if (optcode %in% c("1", "2", "5", "6", "7", "8")) {
      
      input1_index <- switch(
        param1, 
        "0" = running_state[instr_index + 1] + 1, 
        "1" = instr_index + 1
      )
      
      input2_index <- switch(
        param2,
        "0" = running_state[instr_index + 2] + 1,
        "1" = instr_index + 2
      )
      
      input1 <- running_state[input1_index]
      input2 <- running_state[input2_index]
      
      if (optcode %in% c("1", "2")) {
        
        output <- switch(optcode, "1" = input1 + input2, "2" = input1 * input2)
        
        output_index <- running_state[instr_index + 3] + 1
        running_state[output_index] <- output
        
        instr_index <- instr_index + 4
        
      }
      
      if (optcode == "5") {
        instr_index <- ifelse(input1 == 0, instr_index + 3, input2 + 1)
      }
      
      if (optcode == "6") {
        instr_index <- ifelse(input1 == 0, input2 + 1, instr_index + 3)
      }
      
      if (optcode  == "7") {
        
        output <- 1 * (input1 < input2)
        
        output_index <- running_state[instr_index + 3] + 1
        running_state[output_index] <- output
        
        instr_index <- instr_index + 4
        
      }
      
      if (optcode  == "8") {
        
        output <- 1 * (input1 == input2)
        
        output_index <- running_state[instr_index + 3] + 1
        running_state[output_index] <- output
        
        instr_index <- instr_index + 4
        
      }
      
    }
    
    if (optcode == "3") {
      
      output_index <- running_state[instr_index + 1] + 1
      running_state[output_index] <- input[1]
      
      input <- input[-1]
      instr_index <- instr_index + 2
      
    }
    
    if (optcode == "4") {
      
      output_index <- running_state[instr_index + 1] + 1
      
      # message("Received input:", running_state[output_index])
      
      output_signal <- running_state[output_index]
      
      instr_index <- instr_index + 2
      
    }
    
    if (optcode == "9") break
  
  }
  
  output_signal
  
}

thruster_signal <- vector("double", length = nrow(posib))

for (r in 1:nrow(posib)) {
  
  signal <- 0
  for (c in 1:ncol(posib)) {
    signal <- run_intcode(phase_setting = posib[r, c], input_signal = signal)
  }
  
  thruster_signal[[r]] <- signal
}

max(thruster_signal)
```


## Part 2

Idea below, but not working and don't have time to figure it out. 

> **Hint:** When I try to run the example `3,26,1001,26,-4,26,3,27,1002,27,2,27,1,27,26,
27,4,27,1001,28,-1,28,1005,28,6,99,0,0,5` with the phase sequence `9,8,7,6,5`, the intcode program fails because it arrives at optcode `3` (the one that asks for the original input) twice, and the second time the `input` vector is empty. Should the original input be recycled for the same phase setting? Or is there something off in my intcode program because optcode `3` should only be encountered once?

```{r, eval=FALSE}
posib <- gtools::permutations(
  n = 5, r = 5, v = 5:9, 
  repeats.allowed=FALSE
  )

thruster_signal <- vector("double", length = nrow(posib))

for (r in 1:nrow(posib)) {
  
  signal <- 0
  c <- 1

  repeat {
    signal <- run_intcode(phase_setting = posib[r, c], input_signal = signal)
    c <- ifelse(c < 5, c + 1, 1) 
  }
  
  thruster_signal[[r]] <- signal
}

max(thruster_signal)
```

